Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-10-13 boutiquespaper.tex
more minor edits (done for the day)
updated results, discussion, and conclusion

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
% gk: I agree with not including the above
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the manifest and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

@@

%%Containers are not mandatory though. If an application manifest does
%%not contain any reference to a container, the platform will assume
%%that the application is pre-installed on the execution infrastructure. 
% gk: I believe we only support containers now, correct?

@@

The following platforms are currently able to integrate and execute Boutiques
applications.

@@

CBRAIN~\cite{SHER-14} is a web platform that can process data
distributed into multiple storage locations on computing clusters and
clouds. CBRAIN offers transparent access to remote data sources,
distributed computing sites, and an array of processing and
visualization tools within a controlled, secure environment.  The
CBRAIN service deployed at the Montreal Neurological Institute relies
on the infrastructure provided by Compute Canada~\cite{das2016mni}. It
currently provides 460+ collaborators in 20 countries with web access
to several systems, including 6 clusters of the Compute Canada
high-performance computing infrastructure (totaling more than 100,000
computing cores and 40PB of disk storage) and Amazon EC2. CBRAIN
transiently stores about 10 million files representing over 50TB
distributed in 42 servers. 56 data processing applications are integrated and
over 340,000 processing batches have been submitted since 2010.

Applications in CBRAIN are integrated as Ruby classes that create web
forms, validate parameters and run command lines on computing
resources. \boutiques is supported through a set of templates that
generate such classes from the application manifest. Two application
integration modes are available:
\begin{enumerate}
  \item The manifest is stored in a CBRAIN plugin and the Ruby classes
    are generated on-the-fly when CBRAIN starts. This mode allows
    CBRAIN developers update all Boutiques applications at once by
    editing the templates. However, it does not allow for
    customization beyond the Boutiques schema.
  \item Ruby classes are generated from manifests
    through an offline process and integrated in CBRAIN as any other
    application. This mode allows developers to customize applications
    by editing the generated Ruby classes, but the resulting
    applications are difficult to maintain in the long term, in
    particular when the manifests are updated.
\end{enumerate}

A custom property (\texttt{cbrain:inherits-from-class}) was added to
the Boutiques manifest to define the Ruby class that should be used as
parent class for the application in CBRAIN. It provides more
flexibility to the first integration model described previously.

CBRAIN was also extended to support the parallelization mechanism
available in Boutiques: the CBRAIN execution agent periodically scans
the execution directories of the parallelized applications and it submits
the invocations defined in the JSON objects written by the
application in this directory.

A new list mechanism was also introduced in CBRAIN to facilitate the
iteration of Boutiques applications on large sets of files. CBRAIN
lists are specific files that contain references to other CBRAIN
files. When a list is passed to a Boutiques application, the elements
in the list are either concatenated in a single command line (when the
corresponding Boutiques input is a list), or a new command line is
generated for every element in the list (when the input is not a
list). Supporting lists as a specific CBRAIN file type allows for
improved validation. For instance, lists that contain references to
non-existent or deleted files can be detected. It also allows users to
edit lists using their own tools such as scripts or spreadsheet
applications.

\subsubsection{Flywheel}

\todo{Gunnar, Nathanael, John}

\subsubsection{Pegasus}

\todo{Rafael}

\subsubsection{SPINE}

\todo{Charles}

@@

The Virtual Imaging Platform (VIP)~\cite{GLAT-13} is a web portal for medical 
simulation  and image data analysis. VIP allows to make applications 
available as a service, and also to benefit from the resources available 
to the biomed Virtual Organisation (VO) of the European Grid Infrastructure 
\footnote{\url{https://www.egi.eu/}}, supported by approximately 65 computing
sites world-wide offering a total of 130 computing clusters
and 5 PB of storage. The VIP service is currently deployed at the Creatis laboratory
\footnote{\url{https://www.creatis.insa-lyon.fr/site7/en}} and uses the DIRAC 
French service \footnote{\url{https://dirac.france-grilles.fr}} to submit jobs on 
EGI grid and cloud resources.  VIP counts, in December 2016, more than 950 
registered users and a growing number of available applications.

Until recently, applications were manually integrated in VIP as workflows written 
in the Gwendia\cite{MONT-09} language and managed  with the MOTEUR\cite{GLAT-08} workflow engine. 
As of today, Boutiques is supported through a "Boutiques application importer" tool, which 
parses the JSON file and automatically generates the corresponding application workflow
and the wrapper script that handles, among others, the execution of the command line itself. 
The use of an application workflow allows to (i) iterate on input lists and generate 
parallel tasks, as well as to (ii) concatenate multiple tools. The applications used in 
the MICCAI challenges, for example, were followed by other tools allowing to compute 
challenge-specific metrics. Tool concatenation is currently handeled at the importer 
level based on pre-defined workflow templates.

@@

Table~\ref{table:applications} summarizes the applications described
with Boutiques and the main features used.

\paragraph{Niak.} The Niak pre-processing pipeline, executed with the Pipeline System
for Octave and Matlab (PSOM)~\cite{bellec2012pipeline}, was integrated
in CBRAIN through Boutiques. The integration uses the parallelization
mechanism described in Section~\ref{sec:parallelization} so that even
the invocations processing a single subject can be parallelized. It also
allows CBRAIN to leverage the efficient agent model used in PSOM, as
described in~\cite{GLAT-16}. The integration required some work in
PSOM to facilitate its invocation as a non-interactive command-line
application. The resulting CBRAIN plugin is available at
\url{https://github.com/SIMEXP/cbrain-plugins-psom}. It could in
principle be used in other platforms that support the parallelization
property, although this has not been tested.

\paragraph{Nipype.} The Nipype workflow engine~\cite{gorgolewski2011nipype} was
instrumented to export to Boutiques applications described in its
internal Python description language. \todo{Update nipype\_cmd to
  support the latest Boutiques schema and complete this paragraph.}

@@

\paragraph{MICCAI challenges.}
Boutiques was used to integrate 23 pipelines in the VIP platform in
the context of two challenges organized by the MICCAI conference in
2016, related to the segmentation of multiple-sclerosis lesions in MR
images (MSSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/msseg-challenge/overview}})
and of tumor volumes in PET images (PETSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/petseg-challenge/overview}}). The
pipelines were ported to VIP and executed on 205 subjects in a few
weeks only, which would not have been possible without
\boutiques. Some pipelines had to be adjusted manually once ported to
the platform for the following reasons:
\begin{enumerate}
\item A pipeline
  required a GPU, which we enabled through the
  \texttt{nvidia-docker}\footnote{\url{https://github.com/NVIDIA/nvidia-docker}}
  tool not supported in Boutiques by default although it could be a possible extension.
\item A pipeline required more than 10 GB of data dependencies (atlas
  data) which exceeded the maximum size allowed for Docker containers
  in our setup. We solved the issue by installing the data in a
  directory of the host server that we mounted in the container. Such
  platform-specific configurations could not possibly be supported by
  \boutiques.
\item A pipeline wrote more than 10 GB of intermediate data in a
  temporary directory of the container located on a 2 GB partition. We
  solved the issue by mounting a host directory in the temporary directory.
\end{enumerate}
The pipelines did not use much of the expressiveness features offered
by Boutiques since the instructions given to the challengers required
that all the pipeline parameters were hardcoded, to facilitate
execution. Two custom properties (\texttt{vip:miccai-challenger-email}
and \texttt{vip:miccai-challenge-team-name}) were added to the
\boutiques manifest to help post-process results in the specific
context of MICCAI challenges.

\paragraph{ndmg.} The NeuroData MRI to Graphs one-click connectome estimation
pipeline~\cite{kiar2016ndmg}, developed in Python and leveraging FSL, was
exported to \boutiques and is available at \url{https://github.com/neurodata/boutiques-tools}.

\paragraph{Other applications.} Other neuroinformatics applications were ported to CBRAIN using Boutiques:
FSL tools and ICA AROMA are available at
\url{https://github.com/aces/cbrain-plugins-neuro} and some pipelines
of the Human Connectome Project (HCP) are at
\url{https://github.com/big-data-lab-team/cbrain-plugins-hcp}.

\todo{Chris, something on BIDS apps?}

\paragraph{VIP applications.}A few tools were imported in VIP using Boutiques.

The first one is a phase contrast simulator called CreaPhase and developed at Creatis. 
The CreaPhase inputs had certain particularies (some nedeed to be enclosed 
in simple quotes, other were vectors of variable size enclosed in brackets)
that required the post-processing of the wrapper script generated by the
Boutiques importer.

The Odin MRI simulator\footnote{\url{http://od1n.sourceforge.net}} was also imported 
in VIP using Boutiques. Since Odin  may require important amounts of computing resources 
(the common use-case consists in iterating on a large number of inputs files), Odin is 
executed by VIP on the EGI grid resources, which currently do not support Docker. Consequently, 
the Odin executable was exported from the Docker image which allowed for his CentOS compilation
and the Odin wrapper script was modified so that it makes no "Docker run" call.

@@

A few other applications from the FLI (France Life Imaging) consortium were also imported in
VIP using Boutiques. One of them is Ginfizz, a FMRI pipeline allowing to execute multiple 
preprocessing steps and basic analyses such as: (i) realignment (motion registration and correction),
(ii) slice timing (temporal interpolation to correct for the delay of acquisition between slices), 
(iii) coregistration (registering functional and structural data), (iv) normalization (registering 
of functional and anatomical data to a match a template that has to be specified), and (iv)
smoothing (spatial filtering to increase Signal to Noise ratio). The use of Docker containers (\url{https://hub.docker.com/r/fliiam/ginfizz} based on \url{https://hub.docker.com/r/miykael/nipype_level2}) 
greatly simplifies its deployment. The animaN4BiasCorrection tool from the Anima  \footnote{\url{https://github.com/Inria-Visages/Anima-Public/wiki}} project was also made available 
in VIP. The import process was straightforward.

@@

    \hline
    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    FSL tools &&&&&&\\
    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
    ndmg                &                   &       \cellcolor{gray!75}            &                   &          \cellcolor{gray!75}         &      \cellcolor{gray!75}             &                   &                   \\
    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
    \end{tabular}
  \vspace*{0.5cm}
  \begin{tabular}{l|ccc|ccc}
    \hline
    \rowcolor[gray]{0.9}
    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
    \rowcolor[gray]{0.9}
                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
    \hline
    Niak pre-processing             &\cellcolor{gray!75} &            &         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
    FSL tools &&&&&\\
    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
    FSL Nipype interfaces           &                    &            &         &              &&\\
    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &              &&\\
    ndmg                            &\cellcolor{gray!75} &            &         &              &&\\
    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &              &&\\
  \end{tabular} 
  \caption{Applications available in Boutiques with manifest features used.}
  \label{table:applications}
\end{table}
\note{greg}{@tristan g: chris helped me verify that ndmg also works with singularity, but we have not shared
the image anywhere. Would it be worth me publishing that image for this table?}
\section{Discussion}

With Boutiques, developers can port their applications once and execute
them in several platforms. \boutiques removes the technological dependency to a
particular platform and facilitates migration.  Although the
motivating use cases were taken from neuroinformatics, our primary
field of interest, nothing prevents the system from being used in a
variety of other domains.

@@

Nonetheless, a complex manifest schema has a cost for application
developers and platforms, which we address as follows. For developers,
we maintain the set of mandatory manifest properties as small as
possible so that simple applications can be described in a few lines
only (see Listing~\ref{listing:minimal}). For platforms, we aim at
supporting as many features as possible in the local executor so that
only the following steps need to be implemented to support \boutiques
in a platform, regardless of the complexity of the manifest:
\begin{enumerate}
  \item Input entry: generate the interface to allow users to enter
    input values.
  \item Input validation: create a JSON object from the interface,
    validate it against the invocation schema.
  \item Execution: generate and run the command line from the local
    executor and input JSON object.
  \item Output delivery: from the manifest, identify the output files
    and deliver them to the user.
\end{enumerate}
In particular, advanced validation features such as dependencies
between inputs are embedded in the regular JSON validation performed
at step 2, without requiring the platform to support the related
manifest properties.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "name": "echo",
    "tool-version": "1.0",
    "description": "A simple script to test output files",
    "command-line": "echo [PARAM] > output.txt",
    "schema-version": "0.4",
    "inputs": [{
	"id": "param",
	"name": "Parameter",
        "value-key": "[INPUT_FILE]",
	"type": "File"
    }],
    "output-files": [{
	"id": "output_file",
	"name": "Output file",
	"path-template": "output.txt"
    }]

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, it remains unclear how to validate that these
applications actually execute as expected. A testing framework could
be designed and potentially fed by existing frameworks, to address
this issue.

Another limitation is related to the security of containerized
applications. Since containers are usually controlled by application
developers rather than platforms, which is a good thing to reduce
application porting bottlenecks, nothing prevents developers to embed
malicious code in their container at any stage of the process,
possibly after a platform administrator inspected the
container. Containers are currently bulky file archives that are
cumbersome to inspect. Tools need to be developed to allow for an
easier characterization of container contents, for instance through
comparison digests with respect to validated base images.

@@

The following differences are conceptual in the sense that they may
not be easily addressed by CWL or Boutiques developers without deeply
refactoring the frameworks.

>>>>

At runtime, a value is assigned to all the mandatory and some of the
optional inputs. %% Inputs of type ``String'' may contain any string,
%% inputs of type "Number" must contain a string representing a number,
%% inputs of type "File" must contain a string representing a file path
%% (absolute or relative to the execution directory) and inputs of type
%% "Flag" must contain a boolean. 
%% When input is a list, the value contains the concatenation of all the
%%values in the list.
%% gk: I agree with not including the above
Algorithm~\ref{algo:command-line} shows how the command line is
constructed from the manifest and the values entered by the user. It
substitutes all the value keys in the command line, output path
templates and configuration files, and writes the configuration files.
\begin{algorithm}[h!]
\caption{Command-line construction}
\label{algo:command-line}
\begin{algorithmic}
  \State \Comment{Substitute input value keys in output path templates, configuration files and command line.}
  \For{\texttt{input} in \texttt{inputs}}
  \If{\texttt{input} has a \texttt{value-key}}
  \For{\texttt{output} in \texttt{outputs}}
  \State \texttt{stripped\_value} = \texttt{input\_value}
  \If{\texttt{input} type is \texttt{File}}
  \State In \texttt{stripped\_value}, remove  all elements in \texttt{path-template-stripped-extensions}.
  \EndIf
  \State In \texttt{path-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \If{\texttt{output} has a \texttt{file-template}}
  \State In any line of \texttt{file-template}, replace all occurrences of \texttt{value-key} by \texttt{stripped\_value}.
  \EndIf
  \EndFor
  \State Prepend \texttt{command-line-flag} and \texttt{command-line-flag-separator} to \texttt{input\_value}.
  \State In \texttt{command-line}, replace all occurrences of \texttt{value-key} by \texttt{input\_value}.
  \EndIf
  \EndFor

@@

%%Containers are not mandatory though. If an application manifest does
%%not contain any reference to a container, the platform will assume
%%that the application is pre-installed on the execution infrastructure. 
%% gk: I believe we only support containers now, correct?

@@

\boutiques application import and execution is currently supported by several
platforms which aid in execution and tool management. These are enumerated
below.

@@

CBRAIN~\cite{SHER-14} is a web platform that can process data
distributed into multiple storage locations on computing clusters and
clouds. CBRAIN offers transparent access to remote data sources,
distributed computing sites, and an array of processing and
visualization tools within a controlled, secure environment.  The
CBRAIN service deployed at the Montreal Neurological Institute relies
on the infrastructure provided by Compute Canada~\cite{das2016mni}. It
currently provides 500+ collaborators in 22 countries with web access
to several systems, including 6 clusters of the Compute Canada
high-performance computing infrastructure (totaling more than 100,000
computing cores and 40PB of disk storage) and Amazon EC2. CBRAIN
transiently stores about 10 million files representing over 50TB
distributed in 42 servers. 51 public data processing applications are
integrated and over 340,000 processing batches have been submitted since 2010.

Applications in CBRAIN are integrated as Ruby classes that create web
forms, validate parameters and run command lines on computing
resources. \boutiques is supported through a set of templates that
generate such classes from the application manifest. Two application
integration modes are available:
\begin{enumerate}
  \item The manifest is stored in a CBRAIN plugin and the Ruby classes
    are generated on-the-fly when CBRAIN starts. This mode allows
    CBRAIN developers update all \boutiques applications at once by
    editing the templates. However, it does not allow for
    customization beyond the \boutiques schema.
  \item Ruby classes are generated from manifests
    through an offline process and integrated in CBRAIN as any other
    application. This mode allows developers to customize applications
    by editing the generated Ruby classes, but the resulting
    applications are difficult to maintain in the long term, in
    particular when the manifests are updated.
\end{enumerate}
%% gk: I think we can skip this last bit, since there is only one type of
%% \emph{boutiques} integration with cbrain, we needn't mention the other?

A custom property (\texttt{cbrain:inherits-from-class}) was added to
the \boutiques manifest to define the Ruby class that should be used as
parent class for the application in CBRAIN. It provides more
flexibility to the first integration model described previously.

CBRAIN was also extended to support the parallelization mechanism
available in \boutiques: the CBRAIN execution agent periodically scans
the execution directories of the parallelized applications and it submits
the invocations defined in the JSON objects written by the
application in this directory.

A new list mechanism was also introduced in CBRAIN to facilitate the
iteration of \boutiques applications on large sets of files. CBRAIN
lists are specific files that contain references to other CBRAIN
files. When a list is passed to a \boutiques application, the elements
in the list are either concatenated in a single command line (when the
corresponding \boutiques input is a list), or a new command line is
generated for every element in the list (when the input is not a
list). Supporting lists as a specific CBRAIN file type allows for
improved validation. For instance, lists that contain references to
non-existent or deleted files can be detected. It also allows users to
edit lists using their own tools such as scripts or spreadsheet
applications.

\subsubsection{Flywheel}
%% gk: shall we check with them, and if nothing, remove?
\todo{Gunnar, Nathanael, John}

\subsubsection{Pegasus}
%% gk: to what extent does this exist?
\todo{Rafael}

\subsubsection{SPINE}
%% gk: shall we check with them, and if nothing, remove?
\todo{Charles}

@@

The Virtual Imaging Platform (VIP)~\cite{GLAT-13} is a web portal for medical 
simulation  and image data analysis. VIP allows to make applications 
available as a service, and also to benefit from the resources available 
to the biomed Virtual Organisation (VO) of the European Grid Infrastructure 
\footnote{\url{https://www.egi.eu/}}, supported by approximately 65 computing
sites world-wide offering a total of 130 computing clusters and 5 PB of storage.
The VIP service is currently deployed at the Creatis laboratory
\footnote{\url{https://www.creatis.insa-lyon.fr/site7/en}} and uses the DIRAC 
French service \footnote{\url{https://dirac.france-grilles.fr}} to submit jobs on 
EGI grid and cloud resources.  VIP counts, in October 2017, more than 1000 
registered users and a growing number of available applications.
%% gk: I guessed 50 more in the last year, since 1000 is a nice number, is this accurate?

Until recently, applications were manually integrated in VIP as workflows written 
in the Gwendia\cite{MONT-09} language and managed  with the MOTEUR\cite{GLAT-08} workflow engine. 
As of today, Boutiques is supported through a "Boutiques application importer" tool, which 
parses the JSON file and automatically generates the corresponding application workflow
and the wrapper script that handles, among other things, the execution of the command line itself. 
The use of an application workflow allows to (i) iterate on input lists and generate 
parallel tasks, as well as to (ii) concatenate multiple tools. The applications used in 
the MICCAI challenges, for example, were followed by other tools allowing to compute 
challenge-specific metrics. Tool concatenation is currently handeled at the importer 
level based on pre-defined workflow templates.

@@

Table~\ref{table:applications} summarizes the applications described
with \boutiques and the main features used.

\paragraph{Niak.} The Niak pre-processing pipeline, executed with the Pipeline System
for Octave and Matlab (PSOM)~\cite{bellec2012pipeline}, was integrated
in CBRAIN through \boutiques. The integration uses the parallelization
mechanism described in Section~\ref{sec:parallelization} so that even
the invocations processing a single subject can be parallelized. It also
allows CBRAIN to leverage the efficient agent model used in PSOM, as
described in~\cite{GLAT-16}. The integration required some work in
PSOM to facilitate its invocation as a non-interactive command-line
application. The resulting CBRAIN plugin is available at
\url{https://github.com/SIMEXP/cbrain-plugins-psom}. It could in
principle be used in other platforms that support the parallelization
property, although this has not been tested.

\paragraph{Nipype.} The Nipype workflow engine~\cite{gorgolewski2011nipype} was
instrumented to export to \boutiques applications described in its
internal Python description language. \todo{Update nipype\_cmd to
  support the latest \boutiques schema and complete this paragraph.}

@@

\paragraph{MICCAI challenges.}
\boutiques was used to integrate 23 pipelines in the VIP platform in
the context of two challenges organized by the MICCAI conference in
2016, related to the segmentation of multiple-sclerosis lesions in MR
images (MSSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/msseg-challenge/overview}})
and of tumor volumes in PET images (PETSEG
challenge\footnote{\url{https://portal.fli-iam.irisa.fr/petseg-challenge/overview}}). The
pipelines were ported to VIP and executed on 205 subjects in a few
weeks only, which would not have been possible without
\boutiques. Some pipelines had to be adjusted manually once ported to
the platform for the following reasons:
\begin{enumerate}
\item A pipeline
  required a GPU, which we enabled through the
  \texttt{nvidia-docker}\footnote{\url{https://github.com/NVIDIA/nvidia-docker}}
  tool not supported in Boutiques by default although it could be a possible extension.
\item A pipeline required more than 10 GB of data dependencies (atlas
  data) which exceeded the maximum size allowed for Docker containers
  in our setup. We solved the issue by installing the data in a
  directory of the host server that we mounted in the container. Such
  platform-specific configurations could not possibly be supported by
  \boutiques.
\item A pipeline wrote more than 10 GB of intermediate data in a
  temporary directory of the container located on a 2 GB partition. We
  solved the issue by mounting a host directory in the temporary directory.
\end{enumerate}
The pipelines did not use much of the expressiveness features offered
by \boutiques since the instructions given to the challengers required
that all the pipeline parameters were hardcoded, to facilitate
execution. Two custom properties (\texttt{vip:miccai-challenger-email}
and \texttt{vip:miccai-challenge-team-name}) were added to the
\boutiques manifest to help post-process results in the specific
context of MICCAI challenges.

\paragraph{ndmg.} The NeuroData MRI to Graphs one-click connectome estimation
pipeline~\cite{kiar2017comprehensive}, developed in Python and leveraging FSL, was
exported to \boutiques and is available at \url{https://github.com/neurodata/boutiques-tools}.
The ndmg pipeline is deployed in CBRAIN via its \boutiques descriptor, and is available
both through Docker and Singularity container environments.

\paragraph{BIDS apps.} Brain Imaging Data Structure supported
applications~\cite{gorgolewski2017bids}, a recent effort for the adoption of BIDS support
in common neuroimaging pipelines, require a standardized set of input and output parameters
by tools. A \boutiques tool has been developed which generates a manifest for any such
BIDS app, enabling them to be deployed within a broad range of high performance computing
infrastructures, such as CBRAIN.

\paragraph{Other applications.} Other neuroinformatics applications were ported to CBRAIN using \boutiques:
FSL tools and ICA AROMA are available at
\url{https://github.com/aces/cbrain-plugins-neuro} and some pipelines
of the Human Connectome Project (HCP) are at
\url{https://github.com/big-data-lab-team/cbrain-plugins-hcp}.

\paragraph{VIP applications.} Several tools were imported in VIP using \boutiques. 
The first one is a phase contrast simulator called CreaPhase and developed at Creatis. 
The CreaPhase inputs had certain particularies (some nedeed to be enclosed 
in simple quotes, other were vectors of variable size enclosed in brackets)
that required the post-processing of the wrapper script generated by the
Boutiques importer.

The Odin MRI simulator\footnote{\url{http://od1n.sourceforge.net}} was also imported 
in VIP using \boutiques. Since Odin  may require important amounts of computing resources 
(the common use-case consists in iterating on a large number of inputs files), Odin is 
executed by VIP on the EGI grid resources, which currently do not support Docker. Consequently, 
the Odin executable was exported from the Docker image which allowed for his CentOS compilation
and the Odin wrapper script was modified so that it makes no "Docker run" call.

@@

A few other applications from the FLI (France Life Imaging) consortium were also imported in
VIP using \boutiques. One of them is Ginfizz, a FMRI pipeline allowing to execute multiple 
preprocessing steps and basic analyses such as: (i) realignment (motion registration and correction),
(ii) slice timing (temporal interpolation to correct for the delay of acquisition between slices), 
(iii) coregistration (registering functional and structural data), (iv) normalization (registering 
of functional and anatomical data to a match a template that has to be specified), and (iv)
smoothing (spatial filtering to increase Signal to Noise ratio). The use of Docker containers
(\url{https://hub.docker.com/r/fliiam/ginfizz} based on \url{https://hub.docker.com/r/miykael/nipype_level2}) 
greatly simplifies its deployment. The animaN4BiasCorrection tool from the
Anima\footnote{\url{https://github.com/Inria-Visages/Anima-Public/wiki}} project was also made available 
in VIP. The import process was straightforward.

@@

    \hline
    Niak pre-processing              &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\\
    ICA AROMA                        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    FSL tools &&&&&&\\
    \multicolumn{1}{r|}{anat}        &\cellcolor{gray!75}&\cellcolor{gray!75}&                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{bet}         &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}\\      
    \multicolumn{1}{r|}{fast}        &\cellcolor{gray!75}&                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}\\
    \multicolumn{1}{r|}{first}       &                   &                   &\cellcolor{gray!75}&\cellcolor{gray!75}&\cellcolor{gray!75}&                   &                   \\
    FSL Nipype interfaces     \\
    HCP PreFreesurfer                &                   &                   &                   &                   &                   &                   &                   \\
    ndmg                &                   &       \cellcolor{gray!75}            &                   &          \cellcolor{gray!75}         &      \cellcolor{gray!75}             & \cellcolor{gray!75} &                   \\
    MICCAI challenge (23 pipelines)\\&                   &                   &                   &                   &                   &                   & \\
    \end{tabular}
  \vspace*{0.5cm}
  \begin{tabular}{l|ccc|ccc}
    \hline
    \rowcolor[gray]{0.9}
    Application                     &\multicolumn{3}{c|}{Containers}            &  \multicolumn{3}{c}{Other properties}  \\
    \rowcolor[gray]{0.9}
                                    &Docker              & Singularity& Rootfs  & Config. file & Env vars  & Parallelization   \\
    \hline
    Niak pre-processing             &\cellcolor{gray!75} &\cellcolor{gray!75}&         &              &\cellcolor{gray!75}&\cellcolor{gray!75}\\
    ICA AROMA                       &\cellcolor{gray!75} &            &         &              &&\\
    FSL tools &&&&&\\
    \multicolumn{1}{r|}{anat}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{bet}        &                    &            &         &              &&\\
    \multicolumn{1}{r|}{fast}       &                    &            &         &              &&\\
    \multicolumn{1}{r|}{first}      &                    &            &         &              &&\\
    FSL Nipype interfaces           &                    &            &         &              &&\\
    HCP PreFreesurfer               &\cellcolor{gray!75} &            &         &              &&\\
    ndmg                            &\cellcolor{gray!75} &\cellcolor{gray!75}&         &              &&\\
    MICCAI challenge (23 pipelines) &\cellcolor{gray!75} &            &         &              &&\\
  \end{tabular} 
  \caption{Applications available in Boutiques with manifest features used.}
  \label{table:applications}
\end{table}

\section{Discussion}
With \boutiques, developers can port their applications once and execute
them in several platforms. \boutiques removes the technological dependency to a
particular platform and facilitates migration.  Although the
motivating use cases were taken from neuroinformatics, our primary
field of interest, nothing prevents the system from being used in a
variety of other domains.

@@

Nonetheless, a complex manifest schema has a cost for application
developers and platforms, which we address as follows. For developers,
we maintain the set of mandatory manifest properties as small as
possible so that simple applications can be described in a few lines
only (see Listing~\ref{listing:minimal}). For platforms, we aim at
supporting as many features as possible in the executor so that
only the following steps need to be implemented to support \boutiques
in a platform, regardless of the complexity of the manifest:
\begin{enumerate}
  \item Input entry: generate the interface to allow users to enter
    input values.
  \item Input validation: create a JSON object from the interface,
    validate it against the invocation schema.
  \item Execution: generate and run the command line from the local
    executor and input JSON object.
  \item Output delivery: from the manifest, identify the output files
    and deliver them to the user.
\end{enumerate}
In particular, advanced validation features such as dependencies
between inputs are embedded in the regular JSON validation performed
at step 2, without requiring the platform to support the related
manifest properties.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}
{
    "name": "echo",
    "tool-version": "1.0",
    "description": "A simple script to test output files",
    "command-line": "echo [PARAM] > output.txt",
    "schema-version": "0.4",
    "inputs": [{
	"id": "param",
	"name": "Parameter",
        "value-key": "[INPUT_FILE]",
	"type": "File"
    }],
    "output-files": [{
	"id": "output_file",
	"name": "Output file",
	"path-template": "output.txt"
    }]

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, it remains unclear how to validate that these
applications actually execute as expected beyond executing packaged tests
distributed alongside a given tool. A \boutiques-specific testing
framework could be designed and potentially fed by existing frameworks,
to address this issue.

Another limitation is related to the security of containerized
applications. Since containers are usually controlled by application
developers rather than platforms, which is a good thing to reduce
application porting bottlenecks, nothing prevents developers to embed
malicious code in their container at any stage of the process,
possibly after a platform administrator inspected the
container. Containers are currently bulky file archives that are
cumbersome to inspect. Tools need to be developed to allow for an
easier characterization of container contents, for instance through
comparison digests with respect to validated base images. Singularity
containers have reduced security risks as compared to Docker containers,
but the issue of content transparency is still not avoided.

@@

The following differences are conceptual in the sense that they may
not be easily addressed by CWL or \boutiques developers without deeply
refactoring the frameworks.
